---
title    : Signal through the noise
subtitle : Illustrating the effect of noise on a linear regression
author   : Marc T. Henry de Frahan
widgets  : [mathjax]
mode     : selfcontained
framework   : io2012
---
	   
## This educational Shiny app illustrates the effect of noise on a linear model.
		
- Linear regressions are often used to model complex effects.
- Data measurements always have some associated noise.
- It is important to understand how the noise effects the fit quality.

### <b>
We provide an educational app to explore the effect of Gaussian random noise.
### </b>


---


## Instructions and guidelines

The objective is to fit the following expression with a linear model:
$$y_i=x_i + \varepsilon_i \qquad \text{for}~i = 1,\dots,n$$

where $\varepsilon_i =  \mathcal{N}(\mu, \sigma)~\text{for}~i = 1,\dots,n$.

### <b> 
We provide the user with three sliders: 
### </b>
- the number of points $n$ (our measurements);
- the mean of the Gaussian random noise, $\mu$;
- the standard deviation of the Gaussian random noise, $\sigma$.

The linear model is illustrated via a plot and the $R^2$ value is
provided as a measure of the quality of the fit.

---

## With $n=100$, $\mu=0$, and $\sigma = 0.1$, the quality of the fit is good.

```{r echo = F}
# Initialize the data
require(ggplot2)
set.seed(42)
x <- seq(0,1,length=100)
noise <- rnorm(100, mean = 0, sd = 0.1)
y <- x + noise

# put these in a dataframe
x_name <- "x"
y_name <- "y"
df <- data.frame(x,y)
names(df) <- c(x_name,y_name)
```

```{r lm-plot, fig.height = 4, fig.align = 'center', message = F}
fit <- lm(y ~ x, data = df)
r2 <- summary(fit)$r.squared
ggplot(df, aes(x,y)) +
    geom_point() + geom_smooth(method = "lm", se = FALSE)+
    ggtitle(paste("Fit data with linear model, R-squared=", format(r2,digits=2))) +
    theme(plot.title = element_text(size=22))
```

---

## If $\sigma$ is large ($= 1.2$), $R^2$ is low (poor fit).

```{r echo = F}
x <- seq(0,1,length=100)
noise <- rnorm(100, mean = 0, sd = 1.2)
y <- x + noise
# put these in a dataframe
x_name <- "x"
y_name <- "y"
df <- data.frame(x,y)
names(df) <- c(x_name,y_name)
```

```{r lm-plot2, fig.height = 4, fig.align = 'center', message = F}
fit <- lm(y ~ x, data = df)
r2 <- summary(fit)$r.squared
ggplot(df, aes(x,y)) +
    geom_point() + geom_smooth(method = "lm", se = FALSE) +
    ggtitle(paste("Fit data with linear model, R-squared=", format(r2,digits=2))) +
    theme(plot.title = element_text(size=22))
```


